Foundation models that pass disembodied spatial-reasoning benchmarks
(\textit{e.g.\ } SpartQA \citep{mirzaee2021spartqa}, SpatialVLM
\citep{chen2024spatialvlm}, MMSI-Bench \citep{yang2025mmsi}) struggle
once the agent must \emph{actively} acquire information under partial
observability. \tos \citep{zhang2026theoryofspace} formalises this gap by
asking foundation-model agents to (i) decide what to observe next,
(ii) externalise their cognitive map at every turn, and (iii) revise
that map after a covert false-belief perturbation. Across GPT-5.2,
Gemini-3-Pro, Claude-4.5-Sonnet and three open-source vision-language
models, \tos identifies two consistent failure modes:

\begin{itemize}
\item \textbf{Belief drift} (\tos\,\S5.1): a previously correct
prediction has only $\sim$57--67\% probability of remaining correct in
the next turn (vision-world Stability scores).

\item \textbf{Belief inertia} (\tos\,\S5.3): after objects are silently
relocated, vision agents persist with the obsolete coordinates
\emph{despite directly observing} the new layout (positional inertia
35--69\%, orientational inertia 51--69\%).
\end{itemize}

\paragraph{Hypothesis.}
Both failures share a common architectural root: foundation models
maintain a \emph{flat} belief over individual object coordinates, with
no intermediate structural unit at which (a) to anchor temporal
stability or (b) to localise revision when the world changes. The
mammalian brain does not store space this way: place cells
\citep{okeefe1971} and grid cells \citep{hafting2005} provide a
\emph{region-level} code, and animal territoriality
\citep{burt1943,powell2012homerange} organises familiarity, ownership,
and recovery at the level of the \emph{home range}. Recent neural-AI
replications \citep{banino2018grid} show region-level codes emerge
spontaneously even in artificial agents.

\paragraph{Contribution.}
We introduce \tb, a structurally region-level belief representation,
and plug it as a diagnostic baseline into the \tos benchmark. \tb
organises observed objects under room-level latent units, accumulates a
visit-and-recency familiarity signal, and gates per-object updates by a
confidence buffer that requires \emph{two} contradictory observations
before overwriting a stored fact. We:

\begin{enumerate}
\item formalise \tb as an extension of the \tos belief
$\hat B_t \in \mathcal{J}$ (Section~\ref{sec:method});

\item analytically derive the consequence on Stability and
Belief~Inertia (Section~\ref{sec:method}, full proofs in
Appendix~\ref{app:math});

\item implement a Python post-hoc replay tool that ingests the
\tos exploration log and emits a parallel \tb cognitive map per turn,
without retraining or re-rolling the underlying LLM
(Section~\ref{sec:setup});

\item report preliminary smoke-test numbers and the planned full
Qwen-VL-7B runs (Section~\ref{sec:results}--\ref{sec:plan}).
\end{enumerate}

The framing is explicitly diagnostic rather than performance-chasing:
the goal is to expose \emph{which} \tos failure modes are tied to the
flatness of LLM cognitive maps and which are not, by substituting a
structurally different cognitive map into the same probing pipeline.
