\subsection{Setting (recap of \tos)}

\tos instantiates a partially-observable MDP
$\mathcal{M} = \langle\mathcal{S},\mathcal{A},\Omega,T,O,c,H\rangle$ over
procedural multi-room grids of size $N\times M$ on a tree topology of
rooms with doors as boundary cells. The action space is
$\mathcal{A} = \{\textsc{Goto}(j),\textsc{Rotate}(\theta),\textsc{Observe},\textsc{Query}(j),\textsc{Term}\}$;
observations are egocentric within a $90^{\circ}$ FOV, discretised into
five bearing bins and six distance bins; cost is 1 per \textsc{Observe},
2 per \textsc{Query}; horizon $H{=}20$ for active rollouts. A spatial
belief $B_t$ at turn $t$ is any map from history $h_t$ to a
distribution over world states; \tos elicits the modal estimate
$\hat B_t$ as a JSON cognitive map at every turn (Section~\ref{sec:related}).

\subsection{Definition of \tb}

\begin{definition}[Territorial belief]\label{def:tb}
A territorial belief is the tuple
\[
B^{\mathrm{ter}}_t = \big( \{T_k\}_{k=1}^{K_t},\; A_t,\; \mathcal{O}_t,\; c_t \big),
\]
where each $T_k = (\mathcal{O}_k, D_k, v_k, \tau_k)$ is a territory
(one per visited room) carrying its observed objects, door cells, a
visit count, and the last-visit turn; $A_t = (p_a, \phi_a, k_a)$ is
the agent's pose plus current territory id;
$\mathcal{O}_t : \mathrm{name}\to(k,p,\phi)$ is the global object
table; and $c_t : \mathrm{name}\to\N_{\ge 1}$ is the per-object
confidence counter that gates updates.
\end{definition}

The flat \tos cogmap is recovered as a marginalisation of
$B^{\mathrm{ter}}_t$, so \tb is strictly more expressive
($K{+}N$ extra integer scalars) but, crucially, structurally different:
updates are scoped at the territory level.

\subsection{Update rule}\label{sec:update}

For each visible object $i$ with observed $(b_i, d_i, \phi_i^{\mathrm{obs}})$,
we lift the egocentric observation into a global candidate
position by composing with the agent's pose:
\[
p_i^{\mathrm{new}} = p_a + R(\phi_a)\big(d_i \sin b_i,\; d_i \cos b_i\big)^{\!\top},
\]
where $R(\phi_a)$ rotates local-forward to global $+y$. The object
table is then updated by the four-state machine

\begin{equation}\label{eq:fourstate}
(\mathcal{O}_{t+1}, c_{t+1})(i) =
\begin{cases}
\big((k_a, p_i^{\mathrm{new}}, \phi_i^{\mathrm{new}}),\; 1\big) & i \notin \mathcal{O}_t \;\;\text{\textbf{(new)}}\\[2pt]
\big(\mathcal{O}_t(i),\; c_t(i){+}1\big) & \Delta_i \le \epsilon\;\wedge\; k_a = k_i^{\mathrm{old}}\;\;\text{\textbf{(reinforced)}}\\[2pt]
\big(\mathcal{O}_t(i),\; c_t(i){-}1\big) & \text{contradiction}\;\wedge\; c_t(i) > 1\;\;\text{\textbf{(buffered)}}\\[2pt]
\big((k_a, p_i^{\mathrm{new}}, \phi_i^{\mathrm{new}}),\; 1\big) & \text{contradiction}\;\wedge\; c_t(i) \le 1\;\;\text{\textbf{(overwritten)}}
\end{cases}
\end{equation}
where $\Delta_i = \|p_i^{\mathrm{new}} - p_i^{\mathrm{old}}\|_2$ and
"contradiction" means $\Delta_i > \epsilon$ \emph{or}
$k_a \ne k_i^{\mathrm{old}}$. Agent state is updated unconditionally;
the visited territory increments $v_{k_a^{\mathrm{new}}}$ and stamps
$\tau_{k_a^{\mathrm{new}}} = t{+}1$.

The salient property: \emph{two} contradictions are required to
overwrite a record reinforced at least once. Algorithm~\ref{alg:tb}
gives the executable form.

\begin{algorithm}[t]
\caption{\tb update for one observation}\label{alg:tb}
\begin{algorithmic}[1]
\Require object name $i$, observation $(b_i,d_i,\phi_i^{\mathrm{obs}})$, agent state $A_t$, belief $B_t$
\State $p_i^{\mathrm{new}}\gets p_a + R(\phi_a)(d_i\sin b_i, d_i\cos b_i)^\top$
\If{$i \notin \mathcal{O}_t$} \State store $(k_a,p_i^{\mathrm{new}},\phi_i^{\mathrm{new}})$, $c(i){\gets}1$ \Comment{\textbf{new}}
\Else
  \State $\Delta\gets\|p_i^{\mathrm{new}}-p_i^{\mathrm{old}}\|_2$
  \If{$\Delta\le\epsilon \wedge k_a{=}k_i^{\mathrm{old}}$} $c(i){+\!\!=\!}1$ \Comment{\textbf{reinforced}}
  \ElsIf{$c(i)>1$} $c(i){-\!\!=\!}1$ \Comment{\textbf{buffered}}
  \Else \State store $(k_a,p_i^{\mathrm{new}},\phi_i^{\mathrm{new}})$, $c(i){\gets}1$ \Comment{\textbf{overwritten}}
  \EndIf
\EndIf
\State $v_{k_a}{+\!\!=\!}1$, $\tau_{k_a}{\gets}t{+}1$
\end{algorithmic}
\end{algorithm}

\subsection{Analytical predictions on \tos metrics}

\begin{proposition}[Stability]\label{prop:stab}
Under i.i.d.\ Gaussian observation noise $\eta\sim\mathcal{N}(0,\sigma^2 I)$,
a flat-cogmap update that overwrites on every observation has
$\E[\stab_{\mathrm{flat}}] = \tfrac12$. A buffered \tb update with
buffer $b$ has spurious-overwrite probability bounded by
$(1-\mathrm{CDF}_\sigma(\epsilon))^b$.
\end{proposition}

\begin{proposition}[Belief Inertia]\label{prop:inertia}
For an autoregressive flat update with prior weight $\alpha$, the \tos
positional inertia satisfies $s_i^{\mathrm{pos}} \to \alpha$ as
observation noise vanishes. The buffered \tb rule yields a
\emph{bimodal} inertia profile: $s_i^{\mathrm{pos}}{=}1$ on the first
contradicting observation, $s_i^{\mathrm{pos}}{\to}0$ after the second.
For an object whose territory id changes after the perturbation,
\tb overwrites on the first observation, giving $s_i^{\mathrm{pos}}{\to}0$
immediately.
\end{proposition}

Proofs in Appendix~\ref{app:math}.
