\paragraph{Day~1 (today, evening).}
AutoDL RTX-5090 32\,GB; \verb|vllm serve Qwen/Qwen3-VL-7B-Instruct|;
\verb|git clone --branch release| \tos repository;
\verb|source setup.sh| (downloads \verb|MLL-Lab/tos-data|);
\verb|bash tos_extension/run_full_pipeline.sh| with
\verb|RENDER_MODE=text NUM_SCENES=25|. Deliverable: real
text-world bar chart that replaces Figure~\ref{fig:metric}.

\paragraph{Day~2.}
Same pipeline with \verb|RENDER_MODE=vision|. (Vision images come
pre-rendered; no Unity build required.) If time permits, the
false-belief subset is run and Belief~Inertia numbers are computed
per \tos\,\S5.3, on both LLM and \tb. Optional A800 ablation with
Qwen-VL-32B INT4 to demonstrate the \tb advantage is not a
small-model artefact.

\paragraph{Day~3.}
Stretch goals: familiarity-weighted overwrite
$\epsilon_k(t) = \epsilon_0 (1 - f_k(t))$ (Section~\ref{app:math});
a \verb|compare_baseline| variant restricted to false-belief turns
to surface inertia separately. Final figures, polished writeup,
NeurIPS-style submission draft.

\paragraph{Out of scope.} No fine-tuning or RL training (neither
budget nor time); no new environment beyond the \tos procedural
3-room layouts; no multi-agent extension beyond the conclusion.

\paragraph{Risks.}
The principal risks and mitigations are tabulated below.

\begin{center}\small
\begin{tabular}{p{0.45\linewidth}p{0.45\linewidth}}
\toprule
Risk & Mitigation \\\midrule
\verb|vllm serve| fails to load Qwen-VL-7B & fall back to
GPT-4o-mini API (cost $\le$ \$5 for 25 scenes) \\
\tos \verb|setup.sh| cannot reach HF & pre-download
\verb|MLL-Lab/tos-data| on Mac and rsync \\
Real-data \tb advantage smaller than predicted & report honestly;
the diagnostic itself is the contribution \\
Time runs out on Day~3 & scope to text-world Stability;
mark \S5.3 inertia as future work \\
\bottomrule
\end{tabular}
\end{center}
